---
title: "Demovar verification"
author: "Gilbert M."
date: "Created on: 20 April, 2020, Updated on: `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc_float: false
    includes:
      before_body: [include_1921_nav.html]
---


*** 

Various tabulation and verification for demovars.

### Missing Stats
#### Missing summary for all variables

RelationToHead has the highest missing value followed by Surname and age and MaritalStatus. The number of missing for Surname may be reduced by replacing mentions of " with preceding surname.

```{r, include = FALSE}
offpath <- "Z:/Recensements_1852_1921/1921"
MissStat <- readRDS(file.path(offpath,"WorkingDataGilbert/MissStat.rds"))
require(DT)
```


```{r, echo= FALSE}
datatable(MissStat, options = list(
  order=list(list(0,"asc")),
  columnDefs =list(list(className="dt-left",targets=0)),
  scrollX=T
  ),
  rownames = F
)
```

Result above show there are 11380 cases where all variables are empty. A look at the manuscript show that these cases correspond to cover pages. These recordes can be removeved safely from all files.

See an example at https://www.ancestryinstitution.ca/imageviewer/collections/8991/images/1921_001-E002853146 (Need to be logged in with ancestry)

Below is a list of 10 random image file names for further investigation. Just replace the number in the link above, to view its image.

```{r,eval=FALSE}
> sample(MissPage$ImageFileName,10)
 [1] 1921_115-E003071112 1921_130-E003090609 1921_045-E002910325
 [4] 1921_025-E002883750 1921_114-E003069466 1921_063-e002931806
 [7] 1921_141-E003103610 1921_113-E003068923 1921_161-E003223397
[10] 1921_156-E003123247
196235 Levels: ?ImageFileName 1921_001-E002852803 ... 1921_163-E003226528
> 

```


#### Percentage percentage by province

A difference in frequency for a given variable and its CCRI equivalent is the number of mentions without CCRI coding for that variable.

```{r, include = FALSE}
offpath <- "Z:/Recensements_1852_1921/1921"
MissStat <- readRDS(file.path(offpath,"WorkingDataGilbert/MissStat1.rds"))
require(DT)
```


```{r, echo= FALSE}
datatable(MissStat,  filter = "top", options = list(
  order=list(list(0,"asc")),
  columnDefs =list(list(className="dt-left",targets=0)),
  scrollX=T
  ),
  rownames = F
)
```



### Age and Sex
#### Age (group) pyramid by Sex


```{r, include = FALSE}
require(DT)
require(data.table)
require(ggplot2)

offpath <- "Z:/Recensements_1852_1921/1921"
AgeSex <- readRDS(file.path(offpath,"WorkingDataGilbert/AgeSex.rds"))
AgeSex[,`CCRI_Age(ResidenceAge)`:=sub(' \\(.*', '', `CCRI_Age(ResidenceAge)`)]
AgeSex[`CCRI_Age(ResidenceAge)`=="99999001",`CCRI_Age(ResidenceAge)`:=NA]
AgeSex[,`CCRI_Age(ResidenceAge)`:=as.numeric(`CCRI_Age(ResidenceAge)`)]

AgeSex <- AgeSex[,1:3]
#AgeSex <- AgeSex[!is.na(`CCRI_Age(ResidenceAge)`),1:3]
colnames(AgeSex)<- c("Age","Male","Female")
AgeSex <- melt.data.table(AgeSex, 
                value.name='Population', 
                variable.name = 'Gender', 
                id.vars='Age')


## barplots for male populations goes to the left (thus negative sign)
AgeSex$Population <- ifelse(AgeSex$Gender == "Male",
                            -1*AgeSex$Population,
                            AgeSex$Population)
AgeSex$Age <- cut(AgeSex$Age, breaks = seq(0, 100, 5), right = FALSE) 

```


```{r, echo= FALSE}
n1 <- ggplot(AgeSex, aes(x = Age, y = Population, fill = Gender)) + 
  geom_bar(data = subset(AgeSex, Gender == "Female"), stat = "identity") +
  geom_bar(data = subset(AgeSex, Gender == "Male"), stat = "identity") + 
  scale_y_continuous(labels = paste0(as.character(c(seq(2, 0, -1), seq(1, 2, 1))), "m")) + 
  coord_flip() + 
  scale_fill_brewer(palette = "Set1") + 
  theme_bw()

n1
```


#### Age (details) and Sex

```{r, include = FALSE}
offpath <- "Z:/Recensements_1852_1921/1921"
AgeSex <- readRDS(file.path(offpath,"WorkingDataGilbert/AgeSex.rds"))
require(DT)
```


```{r, echo= FALSE}
datatable(AgeSex, options = list(
  order=list(list(0,"asc")),
  columnDefs =list(list(className="dt-left",targets=0)),
  scrollX=T
  ),
  rownames = F
)
```


### Marital Status and Sex


```{r, include = FALSE}
offpath <- "Z:/Recensements_1852_1921/1921"
MarSex <- readRDS(file.path(offpath,"WorkingDataGilbert/MarSex.rds"))
require(DT)
```


```{r, echo= FALSE}
datatable(MarSex, options = list(
  order=list(list(0,"asc")),
  columnDefs =list(list(className="dt-left",targets=0)),
  scrollX=T
  ),
  rownames = F
)
```

### Age and Marital Status

#### Age (group) and Marital Status
```{r, include = FALSE}

offpath <- "Z:/Recensements_1852_1921/1921"
AgeMar <- readRDS(file.path(offpath,"WorkingDataGilbert/AgeMar.rds"))
setnames(AgeMar, old = "CCRI_Age(ResidenceAge)", new = "Age")

AgeMar[,Age:=sub(' \\(.*', '', Age)]
AgeMar[Age=="99999001",Age:=NA]
AgeMar[,Age:=as.numeric(Age)]

AgeMar$Age <- cut(AgeMar$Age,breaks = seq(0, 100, 5), right = FALSE) 

AgeMar <- melt.data.table(AgeMar, 
                          value.name='freq', 
                          variable.name = 'Marstat', 
                          id.vars='Age')

AgeMar <- AgeMar[,.(freq=sum(freq)),by=c("Age","Marstat")]

AgeMar <- dcast.data.table(AgeMar, Age ~ Marstat,value.var = 'freq')


require(DT)
```


```{r, echo= FALSE}
datatable(AgeMar, options = list(
  order=list(list(0,"asc")),
  columnDefs =list(list(className="dt-left",targets=0)),
  scrollX=T
  ),
  rownames = F
)
```



#### Age (details) and Marital Status
```{r, include = FALSE}
offpath <- "Z:/Recensements_1852_1921/1921"
AgeMar <- readRDS(file.path(offpath,"WorkingDataGilbert/AgeMar.rds"))
require(DT)
```


```{r, echo= FALSE}
datatable(AgeMar, options = list(
  order=list(list(0,"asc")),
  columnDefs =list(list(className="dt-left",targets=0)),
  scrollX=T
  ),
  rownames = F
)
```


### Sex by Geography

```{r, include = FALSE}
offpath <- "Z:/Recensements_1852_1921/1921"
SexPro <- readRDS(file.path(offpath,"WorkingDataGilbert/SexPro.rds"))
require(DT)
```


```{r, echo= FALSE}
datatable(SexPro, filter = "top", options = list(
  order=list(list(0,"asc")),
  columnDefs =list(list(className="dt-left",targets=0)),
  scrollX=T
  ),
  rownames = F
)
```





















